---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# exclude

<!-- badges: start -->
<!-- badges: end -->

The goal of `exclude` is to keep track of excluded dataframe rows.
It aims to be non-intrusive and descriptive. Non-intrusive here means that,
for example, the base R `subset()` function or the `dplyr` `filter()` function 
don't need wrappers,
and can be called as always. The exclusions are measured and logged using additional
function calls (`init_exclude()` and `exclude()`). The logging is descriptive in that
user can specify a high-level description of the exclusion to the `exclude()` function
instead of a mechanical description of how the exclusion was done. For example,
the exclusion message can be "exclude minors" instead of "`filter(age >= 18)`".

## Installation

You can install the development version of `exclude` like so:

``` r
devtools::install_github("FRCBS/exclude")
```

## Examples

This is a basic example which shows you how to solve a common problem:

```{r setup}
library(exclude)
library(tidyverse)
```

The `init_exclude()` call measures the number of rows in the input dataframe,
and the `exclude()` function measures it again and computes the difference, and
logs the exclusion (with default message "Exclusion").

```{r Example 1}
## basic example code
filtered_mpg <- mpg %>% 
  init_exclude() %>%
  filter(manufacturer == "audi") %>%
  exclude()
e <- get_exclude()
e
```

The exclusion object can be converted to a tibble:

```{r}
as_tibble(e)
```

You can also plot the object with `plot(e)`, for example:

```{r, eval=FALSE}
plot(e)
```

<!--
Since plot calls DiagrammeR::grViz(), which returns htmlwidget, it cannot be included in an
md file directly. Below I use DiagrammeRsvg::export_svg() to convert htmlwidget to svg,
and then refer to the svg file from the md file. Other solution is to use the webshot2 package:
htmlwidgets::saveWidget(plot(e), "/tmp/x.html")
webshot2::webshot("/tmp/x.html", "/tmp/x.png")
And the refer to the png file from the md file.
-->

```{r plot1, echo=FALSE}
library(DiagrammeRsvg)
cat(export_svg(plot(e)), file = "man/figures/plot.svg")
```

![](man/figures/plot.svg)

Exclusions using base R can also be tracked, but exclude works more elegantly with
pipes.

```{r}
init_exclude(mpg) 
filtered_mpg <- subset(mpg, manufacturer == "audi")
invisible(exclude(filtered_mpg))
e <- get_exclude()
e
```

An example with several tracked values.

```{r Example 2}
## basic example code
statistics <- function(df) { list(count = nrow(df), models=n_distinct(df$model)) }
mpg %>% 
  init_exclude(statistics = statistics) %>%
  filter(manufacturer == "audi") %>%
  exclude("Exclude non-Audi") %>%
  filter(cyl < 8) %>%
  exclude("Exclude 8-cylinder cars")
e2 <- get_exclude()
e2
```

```{r, eval=FALSE}
plot(e2)
```

```{r plot2, echo=FALSE}
cat(export_svg(plot(e2)), file = "man/figures/plot2.svg")
```

![](man/figures/plot2.svg)

In addition to exclusions in a dataframe, exclusions in a vector can be tracked as well.
For that the default statistics has to be modified.

```{r Example 3}
## basic example code
statistics <- function(v) { list(count = length(v)) }   # Measure the length of the vector
v <- mpg$manufacturer
v <- v %>% 
  init_exclude(statistics = statistics) %>%
  discard(~ str_starts(., "h")) %>%
  exclude("Manufacturer starts with and h letter") 
e <- get_exclude()
e
```

Set the option `exclude.print_messages` to `FALSE` to disable exclusion messages:

```{r}
options(exclude.print_messages = FALSE)
```

## Similar packages

[tidylog](https://github.com/elbersb/tidylog)

The package is intrusive: it requires wrappers for each of the `dplyr` and `tidyr` operations.
Does not support base R. Has more abilities than exclude, such as keeping track
added or removed columns. It is not descriptive: instead of logging what has been
filtered, it tells how the filtering was implemented. Last CRAN release
is from 2020-07-03, so it might not support the latest updates in tidyverse.

[ExclusionTable](https://cran.r-project.org/web/packages/ExclusionTable/index.html)

The used method is quite intrusive and specific to function `subset()`.

[prismadiagramR](https://cran.r-project.org/web/packages/prismadiagramR/index.html)

Only remotely similar to exclude. Has ability to produce exclusion plots.

